Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-09 exp1/iswc2013.tex
more on linearity problem
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@547 43ef1190-90fd-4dd7-a769-d0121f02b80e

\subsection{Full factorial experiment}

@@

The main objective of this paper is not to establish that the
optimization is significant, it is merely a simulation. It is to
establish a critical practice of evaluations. It is of critical
importance that we understand why some effects are critical.\todo{add
  table of effects} That
``Implement'' is significant is hardly a surprise, that is what we
worked for. That the main effects ``Union'', ``Optional'' and ``Lang''
are strong is also due to that these are fairly demanding things to
evaluate. Since the delay we inserted affects all kinds of joins, it
is also intuitive that interactions between ``Implement'' and those
that require joins are strong. The strong ``BGPComp:Lang'' and
``Lang:Union'' interactions might be due to that the many more triples
needs to be searched for a language tag in one case. However, the
positive ``Implement:TripleC'' interaction evades such explanations;
in fact, it hints that our optimization may not work as well for
larger databases. This should cause some alarm. 
It is a fundamental property of two-level experiments that it can only
model linear effects. If the effect is truly linear, and the variance
is small, the two levels can be quite close together and still yield a
reasonable estimate for the effect, but the effect will be greater
with greater separation. Additionally if the effect is non-linear, the
choice of levels is critical. In this case, the experimenter should be
concerned that the separation between levels for the sizes of the
data-set is too small and should further investigate. Such reasoning
should be applied to all significant effects.

@@

>>>>

\subsection{Full factorial experiment}\label{sec:full}

@@

The main objective of this paper is not to establish that the
optimization is significant, it is merely a simulation. It is to
establish a critical practice of evaluations. It is of critical
importance that we understand why some effects are critical.\todo{add
  table of effects} That
``Implement'' is significant is hardly a surprise, that is what we
worked for. That the main effects ``Union'', ``Optional'' and ``Lang''
are strong is also due to that these are fairly demanding things to
evaluate. Since the delay we inserted affects all kinds of joins, it
is also intuitive that interactions between ``Implement'' and those
that require joins are strong. The strong ``BGPComp:Lang'' and
``Lang:Union'' interactions might be due to that the many more triples
needs to be searched for a language tag in one case. However, the
positive ``Implement:TripleC'' interaction evades such explanations;
in fact, it hints that our optimization may not work as well for
larger databases. This is a clear indication that the separation
between levels for the sizes of the data-set is too small and should
be investigated further. Such reasoning should be applied to all
significant effects.

@@

It is a fundamental property of two-level experiments that it can only
model linear effects. Usually, it is quite sufficient to identify
significant effects even though the underlying response may be
non-linear. As we saw in section~\ref{sec:full}, the choice of levels
may be critical, and if non-linearity is expected, more than two
levels must be considered, or at the very least, attention must be
paid to the choice of levels.\todo{this is a long discussion, how much
  should be done?}
